\begin{thebibliography}{99}

\bibitem[1]{aldrich1} Aldrich J., (1997) \textit{R. A. Fisher and the Making of Maximum Likelihood 1912 – 1922}, Statistical Science
1997, Vol. 12, No. 3, 162-176.

\bibitem[2]{assel} Asselain B., Mould R. F., (2010) \textit{Methodology of the Cox proportional hazards model},  Journal of Oncology 2010, volume 60, Number 5,  403–409.

\bibitem[3]{biecek1} Biecek P., (2011) \textit{Przewodnik po pakiecie R}, Rozprawa doktorska, Oficyna Wydawnicza GiS, wydanie II.

\bibitem[4]{bott1} Bottou L., (2010) \textit{Large-Scale Machine Learning with Stochastic Gradient Descent}.

\bibitem[5]{bott2} Bottou L., (2012) \textit{Stochastic Gradient Descent Tricks}.

\bibitem[6]{burzyk1} Burzykowski T., (2015?) \textit{Notatki do przedmiotu Biostatystyka}, \texttt{https://e.mini.pw.edu.pl/sites/default/files/biostatystyka.pdf}.

\bibitem[7]{cox0} Cox D. R. (1962) \textit{Renewal Theory. Methuen Monograph on Applied Probability
\& Statistics}, London: Methuen.

\bibitem[8]{cox}  Cox D. R., (1972) \textit{Regression models and life-tables (with discussion)}, Journal of the Royal Statistical Society Series B 34:187-220.. 

\bibitem[9]{finkel}  Finkel J. R., Kleeman A., Manning C. D., (2008), \textit{Efficient, Feature-based, Conditional Random Field Parsing}, Proc. Annual Meeting of the ACL.

\bibitem[10]{fisher1} Fisher R. A., (1912) \textit{An absolute criterion for fitting frequency curves}. 

\bibitem[11]{fisher2} Fisher R. A., (1922) \textit{On the mathematical foundations of theoretical statistics}, Philos. Trans. Roy. Soc. London Ser. A 222 309-368.


\bibitem[12]{gauss1} Gauss C. F., (1809) \textit{Theoria Motus Corporum Coelestium}.

\bibitem[13]{gagol1} Gągolewski M., (2014) \textit{Programowanie w języku R}, Wydawnictwo Naukowe PWN.


\bibitem[14]{hald} Hald A., (1949) \textit{Maximum likelihood estimation of the parameters of a normal distribution which is truncated at a known point}, Skandinavisk Aktuarietidskrift, 119-134.

\bibitem[15]{hutch1} Hutchinson J. B., (1928) \textit{The Application of the "Method of Maximum Likelihood" to the Estimation of Linkage}, Genetics. 1929 Nov; 14(6): 519–537.


\bibitem[16]{kenward1} Kenward M. G., Lesaffre E. and Molenberghs G., (1994) \textit{An Application of Maximum Likelihood and Generalized Estimating Equations to the Analysis of Ordinal Data from a Longitudinal Study with Cases Missing at Random}, Biometrics
Vol. 50, No. 4 (Dec., 1994), pp. 945-953.

\bibitem[17]{kotlowski} Kotłowski W., (2012), Notatki do przedmiotu \textit{Techniki Optymalizacji} prowadzonego na Politechnice Poznańskiej, \\ \texttt{http://www.cs.put.poznan.pl/wkotlowski/teaching/wyklad3b.pdf}

\bibitem[18]{legendre1} Legendre A. M., (1804) \textit{Nouvelles m\`ethods pour la d\`etermination des orbites des com\`etes}.

\bibitem[19]{milewski} Milewski S., (2006) Konspekt do przedmiotu \textit{Metody Numeryczne} prowadzonego na Politechnice Krakowskiej, \\ \texttt{http://l5.pk.edu.pl/images/skrypty/Metody\_numeryczne\_1}

\bibitem[20]{millar1} Millar R. B., (2011) \textit{Maximum Likelihood Estimation and Inference: With Examples in R, SAS and ADMB, chapter 6. Some Widely Used Applications of Maximum Likelihood}, John Wiley \& Sons, Ltd.

\bibitem[21]{niemiro} Niemiro W., (2011) Skrypt do przedmiotu \textit{Statystyka} prowadzonego na Uniwersytecie Warszawskim, \\ \texttt{http://www-users.mat.umk.pl/$\sim$wniem/Statystyka/Statystyka.pdf}

\bibitem[22]{norwe} Norwegian Multicentre Study Group, (1981) \textit{Timolol-induced reduction in
mortality and reinfarction}, The New England  Journal of Medicine; 304: 801-7.


\bibitem[23]{mit0} Panchenko D., (2006), Notatki do otwartego kursu MIT \textit{Statistics for Applications, Lecture 2: Maximum Likelihood Estimators.}, \\
\texttt{http://ocw.mit.edu/courses/mathematics/18-443-statistics-for-applications-fall-2006/}



\bibitem[24]{mit1} Panchenko D., (2006), Notatki do otwartego kursu MIT \textit{Statistics for Applications, Lecture 3: Properties of MLE: consistency, asymptotic normality. Fisher information.}, \\
\texttt{http://ocw.mit.edu/courses/mathematics/18-443-statistics-for-applications-fall-2006/}



\bibitem[25]{programikr} R Core Team, (2013) \textit{R: A language and environment for statistical computing.} R Foundation for Statistical Computing, Wiedeń , ISBN 3-900051-07-0, \texttt{http://www.R-project.org/}.

\bibitem[26]{rydl1} Rydlewski J., (2009) \textit{Estymatory Największej Wiarogodności w Uogólnionych Modelach Regresji Nieliniowej}, Rozprawa doktorska.

\bibitem[27]{sokol} Sokołowski A., (2010) \textit{Jak rozumieć i wykonywać analizę przeżycia} \textit{http://www.statsoft.pl/Portals/0/Downloads/Jak\_rozumiec\_i\_wykonac\_analize\_przezycia.pdf}

\bibitem[28]{views} Statistics Views, (2014) \textit{ "I would like to think of myself as a scientist, who happens largely to specialise in the use of statistics"– An interview with Sir David Cox}. 

\bibitem[29]{ther} Therneau T. M., Grambsch P. M., (2000), \textit{Modeling Survival Data: Extending the Cox Model}, Springer.

\bibitem[30]{ADALINE} Widrow B, (1960), \textit{An adaptive "ADALINE" neuron using chemical "memistors"}, Technical Report No. 1553-2, Stanford University.

\bibitem[31]{widrow2} Widrow B, Stearns S. D., (1985), \textit{Adaptive Signal Processing, Prentice Hall}.

\bibitem[33]{wiki1} Wikipedia, encyklopedia wolnego dostępu \texttt{wikipedia.pl}
 
 \bibitem[34]{sfu1} Woodcock S, (2014), Notatki do otwartego kursu Uniwersytetu Simona Frasera \textit{ECON 837, Lecture 11 Asymptotic Properties of Maximum Likelihood Estimators}, \\ \texttt{http://www.sfu.ca/~swoodcoc/teaching/sp2014/econ837/11.mle.pdf}
 
\bibitem[35]{zieli} Zieliński R., (1990) \textit{Siedem wykładów wprowadzających do statystyki matematycznej}, Warszawa, Wydawnictwo Naukowe PWN.


\end{thebibliography}